---
title: "Mathematical Methods for International Commerce "
subtitle: "<br/>Week 1/2: Basic Algebra"
author: "Iegor Vyshnevskyi, Ph.D."
institute: "Sogang University"
date: "March 7, 2025"
output:
  xaringan::moon_reader:
    #css: [default, metropolis, metropolis-fonts]
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
---

<style>
@media print{
  body, html, .remark-slides-area, .remark-notes-area {
    height: 100% !important;
    width: 100% !important;
    overflow: visible;
    display: inline-block;
    }
</style>

<style type="text/css">
.remark-slide-content {
    font-size: 34px;
    padding: 1em 4em 1em 4em;
}
</style>

<style type="text/css">
.my-one-page-font {
  font-size: 28px;
}
</style>

</style>

<style type="text/css">
.my-one-page-font-table {
  font-size: 24px;
}
</style>


```{r setup, include = FALSE}
library(tidyverse)
library(knitr)

opts_chunk$set(fig.width = 10, 
               message = FALSE, 
               warning = FALSE,
               echo = FALSE)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
#install.packages("xaringanthemer")
library(xaringanthemer)
style_mono_accent(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "500", "550i"),
  code_font_google   = google_font("Fira Mono"),
  colors = c(
  red = "#f34213",
  purple = "#3e2f5b",
  orange = "#ff8811",
  green = "#136f63",
  white = "#FFFFFF"
)
)
```

# Why Does Math Matter in Economics?

- **Mathematics is everywhere** in economics and finance.

- Helps us **analyze markets, optimize decisions, and manage risk**.

- Today, we’ll explore **how banks use algebra** in international finance.  

*Let’s dive in!*

---

# What we need to know for today's class?

1. **Basic algebraic operations** (addition, subtraction, multiplication, division).

2. **Negative numbers** and their use in financial calculations.

3. **Algebraic expressions** and their evaluation.

4. **Solving equations** and **inequalities** for financial decision-making.

5. **Fractions** and their application in loan payments.

---

class: inverse, center, middle

# 1. Real-World Example of Algebra in International Finance

---

# Example: International Loan Risk

**Scenario:** A company in Europe borrows **$1 million** from a U.S. bank.  

**Problem:** The company earns revenue in euros (€), but the loan is in dollars ($). 

**Key Question:** How does exchange rate fluctuation impact the company’s debt?  

**Solution:** Use algebra to analyze the exchange rate risk, loan repayment and other questions.

???

Scenario: Currency Exchange & Loan Risk Management
Imagine you are working for an international bank that provides loans to businesses operating across multiple countries. Your task is to assess the financial risk of a European company that wants to borrow $1 million for expansion. Since they earn revenue in euros (€) but the loan is in U.S. dollars ($), you need to evaluate the exchange rate risk and ensure the company can repay the loan.

To analyze the situation, you'll use several key mathematical concepts:


---

# Step 1: Working with Negative Numbers

## **Currency Depreciation & Loan Impact**  

- The exchange rate was **€1 = $1.10** but dropped to **€1 = $1.05** 

- This means the **euro lost value**, making it **more expensive** for the company to repay the loan.  

## Using Negative Numbers in Calculations
### Change in Exchange Rate  
   $$ 
   \Delta ER = 1.05 - 1.10 = -0.05 
   $$  
   **The negative value indicates a depreciation of the euro.**  

---

# Step 1: Working with Negative Numbers (cont)
## Using Negative Numbers in Calculations (cont)

### Change in Debt Due to Depreciation 
   - Before depreciation:  
     $$ D_{\text{old}} = \frac{1,000,000}{1.10} = 909,091 \text{ EUR} $$
   - After depreciation:  
     $$ D_{\text{new}} = \frac{1,000,000}{1.05} = 952,381 \text{ EUR} $$
   - **Increase in debt** (using subtraction):  
     $$ \Delta D = 909,091 - 952,381 = - 43,290 \text{ EUR} $$  
   - Since the result is **negative** *(old-new)*, this means the company **owes more** after depreciation.  

---

# Step 1: Working with Negative Numbers (cont)
## Using Negative Numbers in Calculations (cont)


### Multiplying Negative Numbers: Predicting Future Losses

- Suppose the **exchange rate continues to drop** at a rate of **-0.05 per year**.  
- If the debt increase last time was **43,290 EUR**, what happens **after 3 years**?  
  $$
  \text{Total Debt Increase} = \Delta D \times (-3)
  $$  
  $$
  = 43,290 \times (-3) = -129,870 \text{ EUR}
  $$  

 **The negative result means the company’s debt grows significantly due to continuous depreciation.**  

 ### Practice: Please calculate the debt increase after 5 years.

---

# Step 1: Working with Negative Numbers (cont)
## Using Negative Numbers in Calculations (cont)

### Dividing Negative Numbers: Recalculating Exchange Rate Risk

- If the euro depreciation of **-0.05 per year** continues, how many years until the exchange rate drops to **€1 = $0.85**?  
- Given the total depreciation needed:  
  $$
  \text{Total Change} = 1.05 - 0.85 = 0.20
  $$  
- Solve for time:  
  $$
  \text{Years} = \frac{\text{Total Change}}{\text{Annual Change}} = \frac{0.20}{-0.05} = -4
  $$  
 
  **Since time cannot be negative, this means in 4 years, the exchange rate could reach €1 = $0.85 if depreciation continues.**  

 ### Practice: Please calculate the time needed for the exchange rate to drop to €1 = $0.75.


---

# Step 2: Understanding algebraic expressions & evaluating numerically 

## The Company’s Debt Formula

The company's debt can be represented by the following **algebraic expression**:  
$$
D = \frac{L}{ER}
$$  

Where:  
- \( D \) = Debt in euros (€)  
- \( L \) = Loan amount in dollars ($)  
- \( ER \) = Exchange rate (€ per $)  

---

# Step 2: Understanding algebraic expressions & evaluating numerically (cont)

## Evaluating the Expression Numerically  

### Case 1: Initial Exchange Rate

If:  
- L = 1,000,000 USD  
- ER = 1.20 

Then:  
$$
D = \frac{1,000,000}{1.20}
$$  
$$
= 833,333 \text{ EUR}
$$  

 ### Practice: Please calculate the debt amount when the exchange rate is \( ER = 0.95 \).

---

# Step 3: Simplifying Expressions by Collecting Like Terms 

## Interest Payments on a Loan  

The bank considers **interest payments** when determining the total amount to be repaid.  

The formula for total repayment is:  
$$
T = P + rP
$$  

Where:  
- \( T \) = Total repayment  
- \( P \) = Principal loan amount  
- \( r \) = Interest rate  

---

# Step 3: Simplifying Expressions by Collecting Like Terms (cont)

## Simplifying the Expression  

If the interest rate is **5%** (**r = 0.05 **), substitute the value into the equation:  

$$
T = P + 0.05P
$$


Now, **factor out \( P \) (collecting like terms)**:  

$$
T = (1 + 0.05)P
$$


$$
T = 1.05P
$$


**This simplified formula helps calculate interest payments efficiently in banking.**  

## Practice: Please calculate the total repayment when the principal loan amount is \( P = 250,000 \).

---

# Step 4: Multiplying Out Brackets

- The company signs a **hedging contract** to manage exchange rate risk:  

$$
C = (D + F)(1 + r)
$$  

Where:  
- \( C \) = Cost of the contract  
- \( D \) = Debt in euros  
- \( F \) = Fixed fee for hedging. Assumed to be financed by the bank (incl. in loan).  
- \( r \) = Interest rate 


---

# Step 4: Multiplying Out Brackets (cont)

## Expanding the Brackets

Using the **distributive property** to expand:  

$$
C = D(1 + r) + F(1 + r)
$$


Expanding each term:  

$$
C = D \times 1 + D \times r + F \times 1 + F \times r
$$


$$
C = D + Dr + F + Fr
$$


---
# Step 4: Multiplying Out Brackets (cont)

## Final Expression: Total Cost Breakdown  

$$
C = (D + F) + (D + F)r
$$


🔹 **First part (D + F)** represents the original debt and fee.  
🔹 **Second part (D + F)r** represents the additional cost due to interest.  

**This expanded formula helps in accurately calculating the total cost, including fees and interest.**  

<style type="text/css">
.smaller-font {
  font-size: 18px;
}
</style>

<div class="smaller-font">
Note: hedging contract includes forward contract (F) to lock in exchange rate.
</div>

## Practice: Please calculate the total cost when the debt is \( D = 1,000,000 \), fee is \( F = 50,000 \), and interest rate is \( r = 0.05 \).

---

# Step 5: Factorizing Expressions

- The bank wants to **simplify the total cost formula** for easier calculations:  

$$
C = D(1 + r) + F(1 + r)
$$ 

**We notice that** \( 1 + r \) **is a common factor in both terms.**  

- Since \( 1 + r \) appears in both terms, we can **factor it out**:  

$$
C = (D + F)(1 + r)
$$

**This simplifies the formula and makes calculations easier when adjusting values.**  

## Practice: Please factorize the expression \( T = P + 0.05P \).

---

# Step 6: Fractions in Loan Payments & Risk Calculation 

## Quarterly Loan Payments 

If a **$1,000,000 loan** is split into **quarterly payments**, each payment is calculated as:  

$$
P = \frac{1,000,000}{4}
$$ 


## Converting Payments to Euros (€)  

If the company **pays in euros**, and the exchange rate is **€1 = $1.10**, the payment in euros is:  
$$
P_{\text{EUR}} = \frac{250,000}{1.10}
$$

Simplifying the fraction:  
$$
P_{\text{EUR}} = \frac{250,000 \div 10}{1.10 \div 10} = \frac{25,000}{0.11} = 227,273  \text{ EUR}
$$

**Each quarterly payment costs €227,273 at an exchange rate of 1.10.**  

---
# Step 6: Fractions in Loan Payments & Risk Calculation (cont)

## Impact of Exchange Rate Drop 

If the exchange rate **drops to 1.05**:  

$$
P_{\text{EUR}} = \frac{250,000}{1.05}
$$

Simplifying:  

$$
P_{\text{EUR}} = \frac{250,000 \times 100}{1.05 \times 100} = \frac{25,000,000}{105} = 238,095  \text{ EUR}
$$

**Now, each quarterly payment costs €238,095.**  

---
# Step 6: Fractions in Loan Payments & Risk Calculation (cont)

## Increase in Payment Due to Exchange Rate Change

$$
\Delta P_{\text{EUR}} = 238,095 - 227,273
$$

$$
= 10,822 \text{ EUR}
$$

**The company must now pay an additional €10,822 per quarter due to currency depreciation.**  

---

# Step 6: Fractions in Loan Payments & Risk Calculation (cont)

## Multiplying Fractions – Adjusting Loan Payments  

If the bank charges a **processing fee of 1/200 (0.5%)** on each quarterly payment, the extra fee per quarter is:  

$$
F = P_{\text{EUR}} \times \frac{1}{200}
$$

$$
F = \frac{238,095}{200}
$$

- Divide numerator and denominator by 5:  
$$
\frac{47,619}{40} = 1,190
$$  

**The additional fee per quarter is €1,190.**  

## Practice: Please calculate the processing fee if the quarterly payment is €200,000.

---

# Step 7: Solving Equations for Loan Repayment 

## Loan Repayment Over 5 Years

The company wants to **repay the loan over 5 years** with equal annual payments.  

The total loan amount is **$1,000,000**, and the **annual payment is \( P \)**.  

### The equation for equal payments 

$$
1,000,000 = 5P
$$

### Solving for \( P \) (Annual Payment)

To isolate \( P \), **divide both sides by 5**:  

$$
P = \frac{1,000,000}{5} = 200,000 \text{ USD}
$$

**Each annual payment is $200,000.**  

## Practice: Please calculate the annual payment if the total loan amount is $500,000 and the repayment period is 3 years.

---

# Step 8: Linear Inequalities in Risk Management


## Recognizing Inequality Symbols

The following symbols are used in **risk management** and **financial decision-making**:  

| Symbol | Meaning | Example |
|--------|---------|---------|
| $$<$$ | Less than | Profit \( < 100,000 \) (profit is below €100K) |
| $$>$$ | Greater than | Revenue \( > 500,000 \) (revenue exceeds €500K) |
| $$\leq$$ | Less than or equal to | Risk score $$\leq 70$$ (acceptable risk level) |
| $$\geq$$ | Greater than or equal to | Capital reserves $$\geq 1,000,000$$ (must maintain reserves of €1M or more) |

---

# Step 8: Linear Inequalities in Risk Management (cont)

## Bank's Risk Threshold Example

The bank considers a loan **risky** if the company's **annual revenue falls below €900,000**.  

Thus, the company's revenue must satisfy the inequality:  

$$
R - C \geq 900,000
$$  

Where:  
- \( R \) = Annual revenue  
- \( C \) = Annual costs  
- \( 900,000 \) = Minimum revenue required to avoid risk  

---
# Step 8: Linear Inequalities in Risk Management (cont)

## Solving the Inequality for Minimum Required Revenue 

If the company has **annual costs of €150,000**, we substitute \( C = 150,000 \):  

$$
R - 150,000 \geq 900,000
$$

To isolate \( R \), **add 150,000 to both sides**:  

$$
R \geq 1,050,000
$$

**The company must earn at least €1.05M per year to remain financially stable.**  

## Practice: Please solve the inequality for a minimum revenue of €1,200,000 if costs are €250,000.

---
# Step 8: Linear Inequalities in Risk Management (cont)

## Graphical Representation of the Inequality 

The inequality $$ R \geq 1,050,000 $$ can be visualized on a number line:  

- **Red region**: Revenue below €1.05M (Risky )  
- **Green region**: Revenue €1.05M or more (Safe)  

$$
\begin{array}{ccccccccccc}
{\color{red} \bullet} & {\color{red} \bullet} & {\color{red} \bullet} & {\color{red} \bullet} & {\color{black} |} & {\color{green} \bullet} & {\color{green} \bullet} & {\color{green} \bullet} & {\color{green} \bullet} & {\color{green} \bullet} & \cdots \\
\cdots & \mathit{900K} & \mathit{950K} & \mathit{1M} &  \mathbf{1.05M} & \mathit{1.1M} & \mathit{1.2M} & \mathit{1.3M} & \cdots \\
\end{array}
$$

**The company must be in the green region to meet the bank's stability requirements.**  


---

# Conclusion: Why This Math Matters

1. Algebra helps predict financial outcomes.  

2. Fractions, equations, and inequalities are tools for banking decisions.  

3. Math makes economic decision-making precise and reliable.

---

# Next Steps

🔹 Practice **algebra problems from the textbook** (Jacques, Sections 1.1, 1.2). 

🔹 Bring any questions to our **next class discussion!**  

**Math is powerful—and fun!** 


---

class: inverse, center, middle

# Any QUESTIONS?

---

# Next Class

- (Mar 12) Graphs of Equations (1.3), Solving Equations (1.4) 




???
1. To print pdf slides
https://stackoverflow.com/questions/54968311/xaringan-export-slides-to-pdf-while-preserving-formatting

pagedown::chrome_print("W1_ME.html") # but not all pictures are visible

2. Option: https://stackoverflow.com/questions/54968311/xaringan-export-slides-to-pdf-while-preserving-formatting

install.packages("remotes")
remotes::install_github("jhelvy/xaringanBuilder")
remotes::install_github("jhelvy/renderthis@v0.0.9")

library(xaringanBuilder)
build_pdf("DVC.html")

3. Option
writeBin(as.raw(c()), "favicon.ico") # create an empty favicon.ico file
install.packages("renderthis")
remotes::install_github('rstudio/chromote')
library(renderthis)

renderthis::to_pdf("W2_ME.html")

getwd()
setwd("C:/Users/Iegor/OneDrive - kdis.ac.kr/Documents/GitHub/Sogang/2025/Spring/Mathematical Economics/Week 1_2")




---

## Economic Scenario

A European company takes a **loan in US dollars**.

* Loan amount:
  ( L = $1{,}000{,}000 )

* The company earns revenue in **euros**.

* Exchange rate at the time of borrowing:
  ( 1 \text{ EUR} = 1.10 \text{ USD} )

---

## Exchange Rate Convention (Important)

We define the exchange rate as:

[
ER = \text{USD per EUR}
]

Example:

* ( ER = 1.10 ) means **1 euro buys 1.10 dollars**

To convert dollars into euros:

[
\text{EUR} = \frac{\text{USD}}{ER}
]

---

## Step 1: Initial Debt in Euros

The loan is in dollars, but the firm measures debt in euros.

[
D = \frac{L}{ER}
]

[
D = \frac{1{,}000{,}000}{1.10} = 909{,}091 \text{ EUR}
]

---

## Exchange Rate Change

One year later, the euro depreciates:

[
1 \text{ EUR} = 1.05 \text{ USD}
]

The exchange rate falls from 1.10 to 1.05.

---

## New Debt Value

[
D_{\text{new}} = \frac{1{,}000{,}000}{1.05} = 952{,}381 \text{ EUR}
]

---

## Increase in Debt Due to Exchange Rate Change

[
\Delta D = D_{\text{new}} - D_{\text{old}}
]

[
\Delta D = 952{,}381 - 909{,}091 = 43{,}290 \text{ EUR}
]

**Economic interpretation:**
Even though the dollar loan did not change, the firm now owes **€43,290 more** because of exchange rate movements.

---

## Step 2: Modeling Exchange Rate Changes

Assume the exchange rate continues to fall at a constant rate:

[
\Delta ER = -0.05 \text{ USD/EUR per year}
]

We model the exchange rate over time as:

[
ER(t) = 1.05 - 0.05t
]

---

## Predicting a Future Exchange Rate

When will the exchange rate reach 0.85 USD/EUR?

[
1.05 - 0.05t = 0.85
]

[
0.05t = 0.20
]

[
t = 4
]

**Answer:** after **4 years**.

---

## Step 3: Future Debt Level

After 3 years:

[
ER(3) = 1.05 - 0.05 \cdot 3 = 0.90
]

Debt in euros:

[
D = \frac{1{,}000{,}000}{0.90} = 1{,}111{,}111 \text{ EUR}
]

---

## Interpreting the Result

Current debt (at ER = 1.05):

[
952{,}381 \text{ EUR}
]

Debt after 3 years:

[
1{,}111{,}111 \text{ EUR}
]

Increase:

[
158{,}730 \text{ EUR}
]

**Conclusion:**
A falling exchange rate significantly increases the euro value of foreign-currency debt.

---

## Step 4: Annual Repayment Equation

The firm repays the loan over **5 years** with equal annual payments.

Let:

* ( A ) = annual payment (USD)

[
5A = 1{,}000{,}000
]

[
A = 200{,}000
]

---

## Quarterly Payments

There are 4 quarters per year.

[
Q = \frac{A}{4} = \frac{200{,}000}{4} = 50{,}000
]

Each quarterly payment is **$50,000**.

---

## Step 5: Processing Fee

The bank charges a fee equal to:

[
\frac{1}{200} \text{ of the euro debt}
]

Using current debt:

[
F = 952{,}381 \times \frac{1}{200}
]

[
F = 4{,}761.905 \text{ EUR}
]

Approximately:

[
F \approx 4{,}762 \text{ EUR}
]

---

## Step 6: Revenue Requirement

Let:

* ( R ) = annual revenue (EUR)
* ( C = 300{,}000 ) = annual costs (EUR)

Profit:

[
\Pi = R - C
]

The firm is safe if profit is at least €900,000:

[
R - 300{,}000 \ge 900{,}000
]

---

## Solving the Inequality

[
R \ge 1{,}200{,}000
]

**Interpretation:**
Annual revenue must be **at least €1.2 million** to safely service the loan.

---

## Solution Set

[
R \in [1{,}200{,}000, \infty)
]

Any revenue below this level makes the loan risky.

---

## Key Takeaways

* Exchange rates affect real debt burdens.
* Algebra helps convert economic stories into equations.
* Signs and units matter.
* Small exchange rate changes can have large financial effects.

---

## Next Steps

Next week we continue with:

* linear functions,
* slopes and economic interpretation,
* graphical analysis.

(Aligned with *Jacques*, Chapters 1.2–1.4)

---

