---
title: "Mathematical Methods for International Commerce "
subtitle: "<br/>Week 11/2: Lagrange Multipliers (5.6)"
author: "Iegor Vyshnevskyi, Ph.D."
institute: "Sogang University, Geppert School of International Studies"
date: "May 14, 2026"
output:
  xaringan::moon_reader:
    #css: $$default, metropolis, metropolis-fonts$$
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
---

<style>
@media print{
  body, html, .remark-slides-area, .remark-notes-area {
    height: 100% !important;
    width: 100% !important;
    overflow: visible;
    display: inline-block;
    }
}
</style>

<style type="text/css">
.remark-slide-content {
    font-size: 34px;
    padding: 1em 4em 1em 4em;
}
</style>

<style type="text/css">
.my-one-page-font {
  font-size: 28px;
}
</style>

<style type="text/css">
.my-one-page-font-table {
  font-size: 24px;
}
</style>

<style>
.tiny { font-size: 60%; }      /* class you can reuse anywhere */
</style>

<style>
.remark-slide-content {
  position: relative;
  z-index: 1;
}

.remark-slide-content::before {
  content: "";
  position: absolute;
  top: 50%;
  left: 50%;
  width: 600px;          /* adjust size */
  height: 600px;
  background-image: url("1. 교장(Seal_Positive).png");  /* place logo file in same folder */
  background-repeat: no-repeat;
  background-position: center;
  background-size: contain;
  opacity: 0.05;         /* watermark transparency */
  transform: translate(-50%, -50%);
  pointer-events: none;
  z-index: 0;
}
</style>


```{r setup, include = FALSE}
library(tidyverse)
library(knitr)

opts_chunk$set(fig.width = 10, 
               message = FALSE, 
               warning = FALSE,
               echo = FALSE)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
#install.packages("xaringanthemer")
library(xaringanthemer)
style_mono_accent(
  base_color = "#851a10",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "500", "550i"),
  code_font_google   = google_font("Fira Mono"),
  colors = c(
  red = "#f34213",
  purple = "#3e2f5b",
  orange = "#ff8811",
  green = "#136f63",
  white = "#FFFFFF"
)
)
```

# Why It Matters in Economics & Finance

- **Resource Allocation:** Optimal use of limited resources

- **Cost Minimization:** Firms aim to minimize costs given production constraints

- **Utility Maximization:** Consumers maximize utility given budget constraints

---

# Learning Objectives

- Apply the Lagrange multiplier method to solve constrained optimization problems.

- Interpret the economic meaning of the Lagrange multiplier as the marginal value of the constraint.

- Solve optimization problems involving Cobb-Douglas production functions with cost constraints.

- Demonstrate that at the optimum, the marginal product per dollar spent is equalized across inputs.

---

# What Are Lagrange Multipliers?

- Lagrange multipliers provide a way to find the **maximum or minimum** of a function subject to a constraint.
- They are particularly useful in economics for problems involving **cost minimization** or **utility maximization**.
- The method involves introducing a new variable (the Lagrange multiplier) to incorporate the constraint into the optimization problem.

- For a function $f(x, y)$ subject to a constraint $h(x, y) = c$, we define the Lagrangian as:

$$
\mathcal{L}(x, y, \lambda) = f(x, y) + \lambda \cdot (c - h(x, y))
$$

- The Lagrange multiplier $\lambda$ can be interpreted as the **rate of change of the objective function** with respect to the constraint.

---

# Lagrange Multipliers - Example

### Problem Statement

- Optimize the objective function:
  $$ f(x, y) = x^2 - 3xy + 12x $$

- Subject to the constraint:
  $$ 2x + 3y = 6 $$

---

# Step 1: Define the Lagrangian

- The Lagrangian function is given by:

$$
\mathcal{L}(x, y, \lambda) = x^2 - 3xy + 12x + \lambda (6 - 2x - 3y)
$$

---

# Step 2: Derive First-Order Conditions

- Compute the partial derivatives of $\mathcal{L}$:

$$
\frac{\partial \mathcal{L}}{\partial x} = 2x - 3y + 12 - 2\lambda
$$

$$
\frac{\partial \mathcal{L}}{\partial y} = -3x - 3\lambda
$$

$$
\frac{\partial \mathcal{L}}{\partial \lambda} = 6 - 2x - 3y
$$

Set each of these derivatives to zero to form the system of equations.

---

# Step 3: Form the System of Equations

1. $2x - 3y - 2\lambda = -12$

2. $-3x - 3\lambda = 0$

3. $2x + 3y = 6$

---

# Step 4: Eliminate Variables

From equation (2):

$$-3x - 3\lambda = 0 \implies x = -\lambda$$

Substitute $x = -\lambda$ into equation (3):

$$2(-\lambda) + 3y = 6 \implies -2\lambda + 3y = 6 \quad (4)$$

Substitute $x = -\lambda$ into equation (1):

$$2(-\lambda) - 3y - 2\lambda = -12 \implies -4\lambda - 3y = -12 \quad (5)$$

Now solve equations (4) and (5).

---

# Step 5: Substitute to Find $y$ and $x$

- From equation (5), substitute $\lambda = 1$:

$$
-4(1) - 3y = -12
$$

$$
-3y = -8 \implies y = \frac{8}{3}
$$

- Substitute $y = \frac{8}{3}$ and $\lambda = 1$ into equation (1):

$$
2x - 3\left(\frac{8}{3}\right) - 2(1) = -12
$$

$$
2x - 8 - 2 = -12 \implies x = -1
$$

---
class: my-one-page-font
# Step 5_1: Check the second order derivative (Optional for now)

*Note*: We don't use Bordered Hessian here because we have only one constraint and the second derivative test is sufficient to classify the stationary point.

To classify the stationary point, use the constraint:

$$
2x + 3y = 6
$$

Solve for $y$:

$$
3y = 6 - 2x
$$

$$
y = 2 - \frac{2}{3}x
$$

Substitute this into the objective function:

$$
f(x, y) = x^2 - 3xy + 12x
$$

$$
f(x) = x^2 - 3x\left(2 - \frac{2}{3}x\right) + 12x
$$
---
class: my-one-page-font
# Step 5_1: Check the second order derivative (Optional for now)

Expand:

$$
f(x) = x^2 - 6x + 2x^2 + 12x
$$

$$
f(x) = 3x^2 + 6x
$$

Now take the second derivative:

$$
f'(x) = 6x + 6
$$

$$
f''(x) = 6
$$

Since:

$$
f''(x) > 0
$$

the stationary point is a **constrained minimum**.

Therefore:

$$
x = -1, \quad y = \frac{8}{3}
$$

gives a **minimum value** of the objective function subject to the constraint.


---

# Step 6: Calculate Optimal/Minimal Value

- Optimal solution:
  - $x = -1$, $y = \frac{8}{3}$, $\lambda = 1$

- Calculate the objective function at the optimal point:

$$
f(-1, \frac{8}{3}) = (-1)^2 - 3(-1)\left(\frac{8}{3}\right) + 12(-1)
$$

$$
= 1 + 8 - 12 = -3
$$

- Stationary value of the objective function under the constraint : **-3**

---

# Step 7: Economic Interpretation

- The Lagrange multiplier $\lambda$ represents the **marginal value of the constraint**.

- In this case, it indicates how much the objective function would change if the constraint were relaxed by one unit.

- A positive $\lambda$ means relaxing the constraint increases the objective value.

- A negative $\lambda$ means relaxing the constraint decreases the objective value.

- The optimal solution $(x, y)$ represents the best allocation of resources given the constraint.

- The Lagrange multiplier $\lambda$ provides insight into the trade-offs involved in the optimization problem, helping to understand the sensitivity of the objective function to changes in the constraint.

- In this case, the optimal solution suggests that the firm should allocate resources in a way that maximizes the objective function while adhering to the constraint.

---

# Lagrange Multipliers: Profit Maximization

A monopolistic producer of two goods $G_1$ and $G_2$ has the following cost function:

$$
TC = 10Q_1 + Q_1Q_2 + 10Q_2
$$

Demand equations:

$$
P_1 = 50 - Q_1 + Q_2 $$ 

$$ P_2 = 30 + 2Q_1 - Q_2
$$

Objective: Maximize profit given the constraint that the total output is 15 units.

---
# Step 1: Formulate the Lagrangian

Objective function (Profit):

$$
\pi = TR - TC
$$

Total Revenue:

$$
TR_1 = (50 - Q_1 + Q_2)Q_1 = 50Q_1 - Q_1^2 + Q_1Q_2
$$

$$
TR_2 = (30 + 2Q_1 - Q_2)Q_2 = 30Q_2 + 2Q_1Q_2 - Q_2^2
$$

Total Revenue:

$$
TR = 50Q_1 - Q_1^2 + 3Q_1Q_2 + 30Q_2 - Q_2^2
$$

Profit function:

$$
\pi = 50Q_1 - Q_1^2 + 3Q_1Q_2 + 30Q_2 - Q_2^2 - (10Q_1 + Q_1Q_2 + 10Q_2)
$$
---
# Step 1: Formulate the Lagrangian (continued)

Simplify:

$$
\pi = 40Q_1 - Q_1^2 + 2Q_1Q_2 + 20Q_2 - Q_2^2
$$


Constraint/Production quota:

$$
Q_1 + Q_2 = 15
$$

Lagrangian:

$$
L(Q_1, Q_2, \lambda) = 40Q_1 - Q_1^2 + 2Q_1Q_2 + 20Q_2 - Q_2^2 + \lambda (15 - Q_1 - Q_2)
$$

---

# Step 2: First-Order Conditions

1. Derivative with respect to $Q_1$:

$$
\frac{\partial L}{\partial Q_1} = 40 - 2Q_1 + 2Q_2 - \lambda = 0
$$

2. Derivative with respect to $Q_2$:

$$
\frac{\partial L}{\partial Q_2} = 20 + 2Q_1 - 2Q_2 - \lambda = 0
$$

3. Derivative with respect to $\lambda$:

$$
15 - Q_1 - Q_2 = 0
$$

---

# Step 3: Solving the System

Rewriting the equations:

1. $40 - 2Q_1 + 2Q_2 - \lambda = 0$

2. $20 + 2Q_1 - 2Q_2 - \lambda = 0$

3. $Q_1 + Q_2 = 15$

**Step 1: Isolate \(\lambda\)**

From equation (1):

$\lambda = 40 - 2Q_1 + 2Q_2$

From equation (2):

$\lambda = 20 + 2Q_1 - 2Q_2$
---

# Step 3: Solving the System (continued)

Setting them equal:

$40 - 2Q_1 + 2Q_2 = 20 + 2Q_1 - 2Q_2$

Combine and simplify:

$20 = 4Q_1 - 4Q_2$

Divide by 4:

$5 = Q_1 - Q_2$

---

# Step 3: Solving the System (continued)


Now, we have two equations:

1. $Q_1 - Q_2 = 5$  
2. $Q_1 + Q_2 = 15$  

Adding the two equations:

$2Q_1 = 20 \implies Q_1 = 10$

Substitute $Q_1 = 10$ in the second equation:

$10 + Q_2 = 15 \implies Q_2 = 5$

---

# Step 4: Calculate $\lambda$

Substituting $Q_1 = 10$ and $Q_2 = 5$:

$\lambda = 40 - 2(10) + 2(5) = 30$

---

# Step 5: Profit Calculation

Substitute $Q_1 = 10$, $Q_2 = 5$ in the profit function:

$$
\pi = 40(10) - (10)^2 + 2(10)(5) + 20(5) - (5)^2
$$

$$
\pi = 400 - 100 + 100 + 100 - 25 = 475
$$

**Optimal Profit:** 475 units.

---

# Economic Interpretation

- The optimal profit is **475 units**.

- The Lagrange multiplier $\lambda = 30$ represents the **marginal increase in profit** for each additional unit of the constraint (production quota).

- If the quota increases by 1 unit, maximum profit increases by approximately 30 units.

---

# Your turn

# Practice Problem 1: Maximizing $2x^2 - xy$


Use Lagrange multipliers to optimise
$$
-2x^2 - xy $$

subject to
$$
x + y = 12
$$

Give economic interpretation of the solution.

---

# Practice Problem 2: Utility Maximization

A consumer’s utility function is given by
$$
U(x_1, x_2) = 2x_1 x_2 + 3x_1$$

where $x_1$ and $x_2$ denote the number of items of two goods $G_1$ and $G_2$ that are bought. 
Each item costs $1 for $G_1$ and $2 for $G_2$. 
Use Lagrange multipliers to find the maximum value of $U$ if the consumer’s income is $83. 
Estimate the new optimal utility if the consumer’s income rises by $1.

---


class: inverse, center, middle

# 2. Home work #2

---

class: my-one-page-font-table

## Homework #2

- **Due Date:** June 11, 2026, end of the day.
- **Submission Format:** Submit your solutions as a single PDF file via the Cyber Campus.
- **Instructions:** 
  - Clearly show all steps and calculations.
  - Include explanations for your answers where applicable.
  - Ensure your submission is neat and well-organized.
  - Bring any questions to the office hours or email me.

- Work on your Home Assignment #2 (Jacques, 10th edition, Chapters 5-9):
  - Chapter 5.1: Exercise 5.1, Problem 5 (p. 411)
  - Chapter 5.2: Exercise 5.2, Problem 7 (p. 427)
  - Chapter 5.3: Exercise 5.3, Practice Problem inside the chapter (p. 432 only)
  - Chapter 5.4: Exercise 5.4, Problem 5 (p. 456)
  - Chapter 5.5: Exercise 5.5, Problems 6 and 7  (p. 467)
  - Chapter 5.6: Exercise 5.6, Problem 4 (p. 479)
  - Chapter 6.1: Exercise 6.1, Problem 3 (p. 506)
  - Chapter 6.2: Exercise 6.2, Problem 6 (p. 521)
  - Chapter 7.1: Exercise 7.1, Problem 5 (p. 553)
  - Chapter 7.2: Exercise 7.2, Problem 6 (p. 572)
  - Chapter 7.3: Exercise 7.3, Problem 5 (p. 584)
  - Chapter 8.1: Exercise 8.1, Problem 6 (p. 615)
  - Chapter 8.2: Exercise 8.2, Problem 3 (p. 626)
  - Chapter 9.1: Exercise 9.1, Problem 4 (p. 655)
  - Chapter 9.2: Exercise 9.2, Problem 3 (p. 670)

Good luck!

---

## Next Classes

- (May 19) Indefinite Integration (6.1) 

- (May 26) Quiz #2
  - Covers materials after Midterm Exam
---

class: inverse, center, middle

# Any QUESTIONS?

## Thank you for your attention and participation!


???
1. To print pdf slides
https://stackoverflow.com/questions/54968311/xaringan-export-slides-to-pdf-while-preserving-formatting

pagedown::chrome_print("W1_ME.html") # but not all pictures are visible

2. Option: https://stackoverflow.com/questions/54968311/xaringan-export-slides-to-pdf-while-preserving-formatting

install.packages("remotes")
remotes::install_github("jhelvy/xaringanBuilder")
remotes::install_github("jhelvy/renderthis@v0.0.9")

library(xaringanBuilder)
build_pdf("DVC.html")

3. Option
writeBin(as.raw(c()), "favicon.ico") # create an empty favicon.ico file
install.packages("renderthis")
remotes::install_github('rstudio/chromote')
library(renderthis)

renderthis::to_pdf("W11_2_ME.html")

getwd()
setwd("C:\\Users\\vyshn\\OneDrive - kdis.ac.kr\\Documents\\GitHub\\Sogang\\2026\\Spring\\Mathematical Methods for International Commerce\\Week 11_2")
