---
title: "Mathematical Methods for International Commerce "
subtitle: "<br/>Week 14/1: Cramer’s Rule (7.3)"
author: "Iegor Vyshnevskyi, Ph.D."
institute: "Sogang University, Geppert School of International Studies"
date: "June 2, 2026"
output:
  xaringan::moon_reader:
    #css: $$default, metropolis, metropolis-fonts$$
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
---

<style>
@media print{
  body, html, .remark-slides-area, .remark-notes-area {
    height: 100% !important;
    width: 100% !important;
    overflow: visible;
    display: inline-block;
    }
}
</style>

<style type="text/css">
.remark-slide-content {
    font-size: 34px;
    padding: 1em 4em 1em 4em;
}
</style>

<style type="text/css">
.my-one-page-font {
  font-size: 28px;
}
</style>

<style type="text/css">
.my-one-page-font-table {
  font-size: 24px;
}
</style>

<style>
.tiny { font-size: 60%; }      /* class you can reuse anywhere */
</style>

<style>
.visual-break {
  width: 45%;
  margin: 0.5em auto 0.7em auto;
  border-top: 2px solid rgba(133, 26, 16, 0.35);
}

.spacious li {
  margin-bottom: 0.3em;
}
</style>

<style>
.remark-slide-content {
  position: relative;
  z-index: 1;
}

.remark-slide-content::before {
  content: "";
  position: absolute;
  top: 50%;
  left: 50%;
  width: 600px;          /* adjust size */
  height: 600px;
  background-image: url("1. 교장(Seal_Positive).png");  /* place logo file in same folder */
  background-repeat: no-repeat;
  background-position: center;
  background-size: contain;
  opacity: 0.05;         /* watermark transparency */
  transform: translate(-50%, -50%);
  pointer-events: none;
  z-index: 0;
}
</style>


```{r setup, include = FALSE}
library(tidyverse)
library(knitr)

opts_chunk$set(fig.width = 10, 
               message = FALSE, 
               warning = FALSE,
               echo = FALSE)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
#install.packages("xaringanthemer")
library(xaringanthemer)
style_mono_accent(
  base_color = "#851a10",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "500", "550i"),
  code_font_google   = google_font("Fira Mono"),
  colors = c(
  red = "#f34213",
  purple = "#3e2f5b",
  orange = "#ff8811",
  green = "#136f63",
  white = "#FFFFFF"
)
)
```

# Agenda

1. Cramer’s Rule (7.3)

2. Group activity: Solving matrixes

3. Homework #2

---

class: inverse, center, middle

# 1. Cramer’s Rule (7.3)

---
## Why It Matters in Economics, Business & Finance

- Many real-world economic systems are governed by **simultaneous equations**.

- Think of equilibrium conditions in macroeconomic models, or supply-demand analysis in trade.

- **Cramer’s Rule** is a neat algebraic technique for solving linear systems using **determinants**, without explicitly computing the inverse matrix.

You'll learn to:
- See when and why Cramer’s Rule is useful

- Use it to solve 2×2 and 3×3 systems

- Apply it to real examples in **macroeconomics** and **international trade**

---

## Recap: Linear System as Matrix Equation

Given a system:
$$a_{11}x + a_{12}y = b_1 \\
a_{21}x + a_{22}y = b_2$$

It can be written as:
$$Ax = b,
\quad\text{where}\quad
A = \begin{bmatrix} a_{11} & a_{12} \\ a_{21} & a_{22} \end{bmatrix},\quad
x = \begin{bmatrix} x \\ y \end{bmatrix},\quad
b = \begin{bmatrix} b_1 \\ b_2 \end{bmatrix}$$

Usually, we solve this using the inverse:
$$x = A^{-1}b$$

> But when $A^{-1}$ is hard to compute, **Cramer’s Rule** offers a shortcut.

---

## Cramer’s Rule: The Concept

Let $A$ be a square matrix with non-zero determinant. Then the solution $x_i$ to the system $Ax = b$ is:
$$x_i = \frac{\det(A_i)}{\det(A)}$$

Where:
- $A_i$: the matrix formed by replacing the *i-th column* of $A$ with the vector $b$

- $\det(A) \ne 0$: system has a unique solution

- Cramer’s Rule works only for square systems (same number of equations and unknowns)

---

## Example 1: Solve Using Cramer’s Rule

Given:
$$4x + 3y = 10 \\
2x + y = 4$$

Step 1: Coefficient matrix:
$$A = \begin{bmatrix} 4 & 3 \\ 2 & 1 \end{bmatrix}, \quad b = \begin{bmatrix} 10 \\ 4 \end{bmatrix}$$

Step 2: Determinant of $A$:
$$\det(A) = (4)(1) - (3)(2) = 4 - 6 = -2$$

---
## Example 1: Solve Using Cramer’s Rule (cont)

Step 3: Determinants with substituted columns:
$$A_1 = \begin{bmatrix} 10 & 3 \\ 4 & 1 \end{bmatrix},\quad \det(A_1) = 10\cdot1 - 3\cdot4 = 10 - 12 = -2$$

$$A_2 = \begin{bmatrix} 4 & 10 \\ 2 & 4 \end{bmatrix},\quad \det(A_2) = 4\cdot4 - 10\cdot2 = 16 - 20 = -4$$

Final Solution:
$$x = \frac{-2}{-2} = 1, \quad y = \frac{-4}{-2} = 2$$

---
## Example 1: Solve Using Cramer’s Rule (cont)


### Quick Check

Verify the solution:

$$
x = 1, \quad y = 2
$$

Substitute into the original system:

$$
4(1) + 3(2) = 10
$$

$$
2(1) + 2 = 4
$$

Both equations hold ✔

---

# Example 2: Solve Using Cramer’s Rule

Solve the system:

$$\begin{aligned}&x_1 + 2x_2 + 3x_3 = 9 \\
&-4x_1 + x_2 + 6x_3 = -9 \\
&2x_1 + 7x_2 + 5x_3 = 13
\end{aligned}$$

Find $x_1$ using **Cramer’s Rule**.

---

## Step 1: Identify Matrices

Coefficient matrix $A$:

$$A = \begin{bmatrix} 1 & 2 & 3 \\ -4 & 1 & 6 \\ 2 & 7 & 5 \end{bmatrix}$$

Right-hand side vector $b$:

$$b = \begin{bmatrix} 9 \\ -9 \\ 13 \end{bmatrix}$$

Replace the **first column** of $A$ with $b$ to form $A_1$:

$$A_1 = \begin{bmatrix} 9 & 2 & 3 \\ -9 & 1 & 6 \\ 13 & 7 & 5 \end{bmatrix}$$

---

## Step 2: Compute Determinants

Expand $\det(A_1)$ along the top row:

$$\det(A_1) = 9 \cdot \begin{vmatrix} 1 & 6 \\ 7 & 5 \end{vmatrix} - 2 \cdot \begin{vmatrix} -9 & 6 \\ 13 & 5 \end{vmatrix} + 3 \cdot \begin{vmatrix} -9 & 1 \\ 13 & 7 \end{vmatrix}$$

$$= 9(-37) - 2(-123) + 3(-76) = -333 + 246 - 228 = -315$$

Recall:
$$
\begin{vmatrix}
a & b \\
c & d
\end{vmatrix}
= ad - bc
$$

---

## Step 3: Compute $\det(A)$

$$\det(A) = 1 \cdot \begin{vmatrix} 1 & 6 \\ 7 & 5 \end{vmatrix} - 2 \cdot \begin{vmatrix} -4 & 6 \\ 2 & 5 \end{vmatrix} + 3 \cdot \begin{vmatrix} -4 & 1 \\ 2 & 7 \end{vmatrix}$$

$$= 1(-37) - 2(-32) + 3(-30) = -37 + 64 - 90 = -63$$

---

## Step 4: Apply Cramer’s Rule

$$x_1 = \frac{\det(A_1)}{\det(A)} = \frac{-315}{-63} = 5$$

- Use same method to find $x_2$, $x_3$ by replacing 2nd or 3rd column of $A$.

---
class: my-one-page-font
## Real-World Example: Static Macroeconomic Model

Consider a simple Keynesian model:
$$Y = C + I + G \\
C = 100 + 0.8Y$$
Solve for $Y$ and $C$ when $I = 50$ and $G = 20$

Step 1: Substitute:
$$Y = (100 + 0.8Y) + 50 + 20 \Rightarrow Y = 170 + 0.8Y$$

$$Y - 0.8Y = 170 \Rightarrow 0.2Y = 170 \Rightarrow Y = 850$$

$$C = 100 + 0.8 \cdot 850 = 780$$

Rearranging:
$$
Y - C = 70
$$

$$
-0.8Y + C = 100
$$

Expressing this in matrix form:
$$A = \begin{bmatrix} 1 & -1 \\ -0.8 & 1 \end{bmatrix}, \quad x = \begin{bmatrix} Y \\ C \end{bmatrix}, \quad b = \begin{bmatrix} 70 \\ 100 \end{bmatrix}$$

---
## Cramer’s Rule Application

Apply Cramer’s Rule to confirm.

---

## Example 2: Two-Country Trade Model

Country A and B produce goods X and Y.
- 2X + 3Y = 20 (A’s resource constraint)
- 3X + Y = 18 (B’s resource constraint)

Solve for quantities of X and Y that satisfy both.

Matrix Form:
$$A = \begin{bmatrix} 2 & 3 \\ 3 & 1 \end{bmatrix},\quad b = \begin{bmatrix} 20 \\ 18 \end{bmatrix}$$

Compute determinants:
$$\det(A) = 2 \cdot 1 - 3 \cdot 3 = 2 - 9 = -7$$

$$A_1 = \begin{bmatrix} 20 & 3 \\ 18 & 1 \end{bmatrix}, \quad \det(A_1) = 20 \cdot 1 - 3 \cdot 18 = 20 - 54 = -34$$
---
## Example 2: Two-Country Trade Model (cont)

$$A_2 = \begin{bmatrix} 2 & 20 \\ 3 & 18 \end{bmatrix}, \quad \det(A_2) = 2 \cdot 18 - 20 \cdot 3 = 36 - 60 = -24$$

Solution:
$$X = \frac{-34}{-7} \approx 4.86, \quad Y = \frac{-24}{-7} \approx 3.43$$

---
class: my-one-page-font
## Practice Problems: Solve using Cramer’s Rule

- Problem 1:
$$3x + 4y = 10 \\
2x + y = 5$$

- Problem 2:
$$\begin{aligned}
2x + 3y - z &= 5 \\
4x - y + 2z &= 6 \\
-3x + 2y + z &= -4
\end{aligned}$$

- Problem 3: In a 3-sector macro model:
$$Y = C + I + G \\
C = a + bY$$
Set $a = 120, b = 0.75, I = 50, G = 30$. Solve for $Y$ and $C$.

- Problem 4: Trade model:
$$4X + 5Y = 40 \\
6X + 2Y = 38$$
Use Cramer’s Rule to find $X$, $Y$.

---

## Summary

- Cramer’s Rule offers a quick solution for **small** systems using determinants

- Works only if $\det(A) \ne 0$

- Particularly useful for **economic models** involving linear constraints

- Cramer’s Rule becomes inefficient for large systems because determinant calculations become complicated

**Limitations**: Only practical for small systems due to cost of computing determinants.

---
class: inverse, center, middle

# 2. Group activity: Solving matrices
---


class: inverse, center, middle

# Surprise!

## No Group Assignment Today

### Instead:

### We will practice together solving matrix problems step-by-step

### Ask questions, make mistakes, and prepare for the quiz/exam calmly :)

---
class: inverse, center, middle

# 3. Home work #2

---

class: my-one-page-font-table

## Homework #2

- **Due Date:** June 11, 2026, before the start of class.
- **Submission Format:** Submit your solutions as **a single file** via the Cyber Campus.
- **Instructions:** 
  - Clearly show all steps and calculations.
  - Include explanations for your answers where applicable.
  - Ensure your submission is neat and well-organized.
  - Bring any questions to the office hours or email me.

- Work on your Home Assignment #2 (Jacques, 10th edition, Chapters 5-9):
  - Chapter 5.1: Exercise 5.1, Problem 5 (p. 411)
  - Chapter 5.2: Exercise 5.2, Problem 7 (p. 427)
  - Chapter 5.3: Exercise 5.3, Practice Problem inside the chapter (p. 432 only)
  - Chapter 5.4: Exercise 5.4, Problem 5 (p. 456)
  - Chapter 5.5: Exercise 5.5, Problems 6 and 7  (p. 467)
  - Chapter 5.6: Exercise 5.6, Problem 4 (p. 479)
  - Chapter 6.1: Exercise 6.1, Problem 3 (p. 506)
  - Chapter 6.2: Exercise 6.2, Problem 6 (p. 521)
  - Chapter 7.1: Exercise 7.1, Problem 5 (p. 553)
  - Chapter 7.2: Exercise 7.2, Problem 6 (p. 572)
  - Chapter 7.3: Exercise 7.3, Problem 5 (p. 584)
  - Chapter 8.1: Exercise 8.1, Problem 6 (p. 615)
  - Chapter 8.2: Exercise 8.2, Problem 3 (p. 626)
  - Chapter 9.1: Exercise 9.1, Problem 4 (p. 655)
  - Chapter 9.2: Exercise 9.2, Problem 3 (p. 670)

Good luck!




---
## Next Class

- (June 4) Linear Programming (8.1) 


---

class: inverse, center, middle

# Any QUESTIONS?

## Thank you for your attention and participation!


???
1. To print pdf slides
https://stackoverflow.com/questions/54968311/xaringan-export-slides-to-pdf-while-preserving-formatting

pagedown::chrome_print("W1_ME.html") # but not all pictures are visible

2. Option: https://stackoverflow.com/questions/54968311/xaringan-export-slides-to-pdf-while-preserving-formatting

install.packages("remotes")
remotes::install_github("jhelvy/xaringanBuilder")
remotes::install_github("jhelvy/renderthis@v0.0.9")

library(xaringanBuilder)
build_pdf("DVC.html")

3. Option
writeBin(as.raw(c()), "favicon.ico") # create an empty favicon.ico file
install.packages("renderthis")
remotes::install_github('rstudio/chromote')
library(renderthis)

renderthis::to_pdf("W14_1_ME.html")

getwd()
setwd("C:\\Users\\vyshn\\OneDrive - kdis.ac.kr\\Documents\\GitHub\\Sogang\\2026\\Spring\\Mathematical Methods for International Commerce\\Week 14_1")
